---
title: "Reproducible Research: Peer Assessment 1"
author: "Jhonattan Reales"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("lubridate")) install.packages("lubridate")
if (!require("ggplot2")) install.packages("ggplot2")
library(tidyverse)
library(lubridate)
library(ggplot2)
```

## Loading and preprocessing the data

```{r}
data <- read.csv("C:/Users/Uninorte/Dropbox/RESUMEN CURSOS/COURSERA/Reproducible Research/WEEK2/1st project/activity.csv")
data$date <- as_datetime(data$date)
glimpse(data)
```


## What is mean total number of steps taken per day?

I will group by the dates and sum all the steps values omitting the missing values

```{r}
total_steps <- data %>% group_by(date) %>% summarise(steps_day=sum(steps, na.rm = TRUE))
head(arrange(total_steps, desc(steps_day)),10)
total_steps %>% ggplot(aes(x=date,y=steps_day)) + geom_bar(color="black",
                                            stat = "identity",fill="skyblue")+
      ylab("Total steps")+ggtitle("Total steps per day")+theme_bw()+
      theme(plot.title =  element_text(hjust=0.5,size=18))
      
```

#### **Mean - omitting missing values**
```{r}
mean(total_steps$steps_day)
```

#### **Median - omitting missing values**
```{r}
median(total_steps$steps_day)
```


## What is the average daily activity pattern?
```{r}
daily_pattern <- data %>% group_by(interval) %>% summarise(avg=mean(steps, na.rm = TRUE))
daily_pattern %>% ggplot(aes(x=interval,y=avg))+geom_line(size=0.8, color="darkblue")+theme_bw()+
      ylab("Average steps")+ggtitle("Average steps per interval")+
      theme(plot.title =  element_text(hjust=0.5,size=18))

```

```{r}
head(arrange(daily_pattern, desc(avg)),20)

```

**The interval 835 has the maximum average steps across all the days**

## Imputing missing values
```{r}
sum(is.na(data$steps)) #2304 missing steps values
sum(is.na(data$date))
sum(is.na(data$interval))
```

The dataset has 2304 missing steps values.

I will impute the NA's with the average value for the specific time interval, in the specific week day. For example, a NA value for a Thursday at interval 720, will have the average value of all the  Thursday at interval 720.

```{r warning=FALSE}
data$week_day <- weekdays(data$date)
data_for_imputing <- data %>% group_by(week_day,interval) %>% summarise(avg=round(mean(steps, na.rm = TRUE),0))

missing <- data %>% filter(is.na(steps)) %>%
      left_join(data_for_imputing, by=c("week_day","interval")) %>% 
      mutate(steps=avg) %>% select(1:4)

missing$steps <- as.integer(missing$steps)
sum(is.na(missing$steps))

```
The new dataset wouldnt have any missing value:

```{r}
data <- data %>% filter(!is.na(steps))
data <- rbind(data,missing)
data <- arrange(data, date)
data[sample(1:nrow(data))[1:10],]
```
And now, the total steps per day without missing values will looks like this:

```{r}
total_steps2 <- data %>% group_by(date) %>% summarise(steps_day=sum(steps, na.rm = TRUE))
head(arrange(total_steps2, desc(steps_day)),10)
total_steps2 %>% ggplot(aes(x=date,y=steps_day)) + geom_bar(color="black",
                                                           stat = "identity",fill="skyblue")+
      ylab("Total steps")+ggtitle("Total steps per day - Imputing missing data")+theme_bw()+
      theme(plot.title =  element_text(hjust=0.5,size=18))
```


#### **Mean - imputing the missing values**
```{r}
mean(total_steps2$steps_day)
```

#### **Median - imputing the missing values**
```{r}
median(total_steps2$steps_day)
```



## Are there differences in activity patterns between weekdays and weekends?
```{r}
set.seed(2)
data$date_type <- if_else(data$week_day %in% c("sábado","domingo"), "weekend","weekday")
data[sample(1:nrow(data))[1:10],]
```


```{r}
daily_pattern2 <- data %>% group_by(interval,date_type) %>% summarise(avg=mean(steps))
daily_pattern2 %>% ggplot(aes(x=interval,y=avg))+geom_line(size=0.8, color="darkblue")+theme_bw()+
      ylab("Average steps")+ggtitle("Average steps per interval - per day of the week")+
      theme(plot.title =  element_text(hjust=0.5,size=18))+facet_wrap("date_type")

```

Yes, there is a 


















